\chapter*{Preface}
This document includes information about HVF user and developer interfaces,
as well as some internal details.

Chapter~\ref{chap:directory} describes the handling of the directory.
Including the syntax of the directory file.

Chapter~\ref{chap:commands} describes each CP command in detail.  Each
command includes a syntax diagram, as well as other information such as
purpose, notes, and examples.

Chapter~\ref{chap:prog} describes the interfaces a developer can use to
communicate with the hypervisor.

\cbstart
Chapter~\ref{chap:nss} describes the design of the Named Saved Segment (NSS)
feature.  Additionally, it also describes the NSS supplied with HVF.
\cbend

Chapter~\ref{chap:abi} describes HVF's internal ABI.

Appendix~\ref{chap:license} includes a copy of the GNU General Public
License (version 2) under which HVF is licensed.

\cbstart
\section*{Summary of Changes in Second Edition}
Added a chapter describing the proposed NSS support (chapter~\ref{chap:nss}).

Expanded the Register Assignment section to detail the differences between
C~code and assembly, and expanded the table to document floating point
registers (section~\ref{sec:registers}).

Added a section describing the s390 and s390x ABI Stack Frame Layouts
(section~\ref{sec:stackframe}).

Expanded ELF file format description.  It now includes a verbal description
of the ELF file format.  Additionally, the section header has been defined
and described in detail (section~\ref{sec:elf}).

Fixed various spelling and clarity issues.
\cbend
